Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-03-18 draftfinalreport_v2.Rnw
revised draft

Previous measles analyses, including two in New Zealand by Prof. Roberts, estimated the interruption of measles virus transmission can be achieved by herd immunity when approximately 95 percent of the population is homogeneously immune to measles \citep{roberts0,roberts4}. Thus, while New Zealand immunisation activities have led to measles outbreaks becoming less frequent, with decreasing numbers of cases, outbreaks still occur (as described above). Current estimates suggest that approximately 85 to 90 percent of the population is immune to measles (see \autoref{sub:immunestat}), thus the reasons for the ongoing outbreaks are likely due to overall population immunity being less than 95 percent and there being pockets of susceptible, non-immune population remaining. Since 2009, all the outbreaks in New Zealand have been linked to infections acquired (imported) from overseas, though previous work suggests these outbreaks still largely affect school-aged children and children under two years of age. Those under two year olds are thought be be consistently among the most affected age groups because the first of two doses of measles, mumps and rubella vaccine (MMR) is not due until fifteen months.

@@

\begin{figure}
     \centering
     \includegraphics[width=0.9\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence in New Zealand from 1997 to 2014}
     \label{fig:incidence1997}
\end{figure}

%%%%%%%%%%%%%%%%%%
\subsection{Risk of measles infection in New Zealand analyses methods}
\label{sub:regression}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to adjust for confounding due to a lack of independence among risk factors. Specifically $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand were required to adjust for confounding and test whether interactions among covariates provide additional information on risk in demographically--defined subgroups over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of measles infection risk.

@@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 15,height=12>>=
par(mfrow=c(1,1))
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="Frequency",cex.lab=2)
hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=2)
@
\end{center}
\caption{Numbers and age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

@@

@@

\vspace{5mm} %5mm vertical space
\begin{table}
<<label=regmodel,include=FALSE,results=tex,echo=FALSE>>=
library(xtable)
xtb<-xtable(summary(model3))
print(xtb,floating = FALSE)
@
\caption{Summary of the regression model results for measles risk factors for cases from 2007-2014}
\label{table:regression_riskfactors}
\end{table}
\vspace{5mm} %5mm vertical space

>>>>

Previous measles analyses, including two in New Zealand by Prof. Roberts, estimated the interruption of measles virus transmission can be achieved by herd immunity when approximately 95 percent of the population is homogeneously immune to measles %\citep{roberts0,roberts4}.
Thus, while New Zealand immunisation activities have led to measles outbreaks becoming less frequent, with decreasing numbers of cases, outbreaks still occur (as described above). Current estimates suggest that approximately 85 to 90 percent of the population is immune to measles (see \autoref{sub:immunestat}), thus the reasons for the ongoing outbreaks are likely due to overall population immunity being less than 95 percent and there being pockets of susceptible, non-immune population remaining. Since 2009, all the outbreaks in New Zealand have been linked to infections acquired (imported) from overseas, though previous work suggests these outbreaks still largely affect school-aged children and children under two years of age. Those under two year olds are thought be be consistently among the most affected age groups because the first of two doses of measles, mumps and rubella vaccine (MMR) is not due until fifteen months.

@@

\begin{figure}
     %\centering
     \includegraphics[width=0.9\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence in New Zealand from 1997 to 2014}
     \label{fig:incidence1997}
\end{figure}

\subsection{Risk of measles infection in New Zealand analyses methods}
\label{sub:regression}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to adjust for confounding due to a lack of independence among risk factors. Specifically $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand were required to adjust for confounding and test whether interactions among covariates provide additional information on risk in demographically--defined subgroups over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of measles infection risk.

@@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mfrow=c(1,1))
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="Frequency",cex.lab=2)
% hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
% legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=2)
% @
% \end{center}
% \caption{Numbers and age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
% \label{fig:ageinyears}
% \end{figure}

@@

<<echo=FALSE,include=FALSE,results=hide>>=

## get data
# set wd
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mfrow=c(1,1))
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="Frequency",cex.lab=2)
% hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
% legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=2)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
## plot cases / age class
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)

require(reshape2)  # this is the library that lets one flatten out data
require(ggplot2)   # plotting library
popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))

## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
colnames(naive)<-c("AgeInYears","Naive","Cases")

testv$VC<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,0,
ifelse(testv$Dose1Mths >= 0 & is.na(testv$Dose2Mths) == T,1,2))
AgeVac<-table(testv$VC,testv$AgeInYears)

row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)
AgeInYears<-(as.numeric(rownames(AgeVac)))
AgeVac<-cbind(AgeVac,AgeInYears)       
AgeInYears<-(as.numeric(rownames(pop)))
pop<-cbind(pop,AgeInYears)
colnames(pop)<-c("Population","AgeInYears")
AgeV<-merge(pop,AgeVac,by="AgeInYears",all=T)

## regression analyses - measles
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
# test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
#tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

#tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
#"5"="1-5","6"="6-10","7"="6-10","8"="6-10",
#"9"="6-10","10"="6-10"));

tnzd<-revalue(testtable$NZDep, c("1"="1-3","2"="1-3","3"="1-3","4"="4-7",
                                 "5"="4-7","6"="4-7","7"="4-7","8"="8-10",
                                 "9"="8-10","10"="8-10"));

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]

# This chunk uses "results=tex"
library(Hmisc)
latex(ttyr, file="",table.env=FALSE,rowname=NULL)

### denominator data
# setwd("~/Massey_2014/measles/data")
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

#tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
#"5"="1-5","6"="6-10","7"="6-10","8"="6-10",
#"9"="6-10","10"="6-10"));

tnzd<-revalue(testtable$NZDep, c("1"="1-3","2"="1-3","3"="1-3","4"="4-7",
                                 "5"="4-7","6"="4-7","7"="4-7","8"="8-10",
                                 "9"="8-10","10"="8-10"));

## here's the additional fudge the MoH asked for.
## add 5 year olds to next class
dm[11:15,3:12]<-dm[11:15,3:12]+round(dm[6:10,3:12]/3,0)
dm[6:10,3:12]<-round(dm[6:10,3:12]*2/3,0)
## add 2 year olds to next class
dm[6:10,3:12]<-dm[6:10,3:12]+round(dm[1:5,3:12]/3,0)
dm[1:5,3:12]<-round(dm[1:5,3:12]*2/3,0)
## 
# rename
dm$Age<-revalue(dm$Age, c("0-2"="0-1", "3-5"="2-4","6-17"="5-17"));

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
popn
#pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
#"Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
#"Dep9"="6-10","Dep10"="6-10"));

pnzd<-revalue(popn$NZDep, c("Dep1"="1-3","Dep2"="1-3","Dep3"="1-3","Dep4"="4-7",
                            "Dep5"="4-7","Dep6"="4-7","Dep7"="4-7","Dep8"="8-10",
                            "Dep9"="8-10","Dep10"="8-10"));

tp$cases <- 0
cases <- matrix(0, length(tp$merge),1)
rownames(cases) <- tp$merge
cases[ttyr$merge,] <- ttyr$Cases
tp$cases <- cases
PerCapita<-(tp$cases/tp$Popn)
PerCap<-cbind(tp[,c(1:4,6)],round(PerCapita*10000,4))
colnames(PerCap)<-c("NZDep","Age","Ethnicity","Population","Cases","Per capita")

#par(mar=c(12,6,4,2)+0.1)
#par(cex.axis=2)
#plot(tp$cases/tp$Popn*10000,main="",
#        ylab="Per Capita Rate Per 10,000",
#        pch=16,cex=2,col=c(rep(1,10),rep(2,10),rep(3,10),rep(4,10),rep(6,10)),
#xlab="",cex.lab=2,xaxt="n")
#axis(side = 1, at = 1:75, labels=rep(c("0-2yr, 1-3","0-2yr, 6-10","3-5yr, 1-5","3-5yr, 6-10","6-17yr, 1-5","6-17yr, 6-10","18-24yr, 1-5",'18-24yr, 1-5',"25+yr, 1-5","25+yr, 6-10"),5),
#     las=2,cex.axis=1.5)
#mtext(side=1,line=10,text="Age and NZDep",cex=2)
#legend("topleft",c("Asian","European","Maori","MLA","Pacific"),
#col=c(1:4,6),pch=16,bty="n",cex=1.8)
#par(mar = c(5,4,4,2) + 0.1)
#

tp$perCap<-with(tp,tp$cases/tp$Popn*10000)
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(NZDep, cases, fill=Age)) + 
%   geom_bar(stat="identity", position="dodge")+
%  # theme(text=element_text(size=45))+
%   ylab("cases")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(NZDep, cases, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
% #  theme(text=element_text(size=45))+
%   ylab("cases")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(Age, cases, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
% #  theme(text=element_text(size=45))+
%   ylab("cases")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(NZDep, perCap, fill=Age)) + 
%   geom_bar(stat="identity", position="dodge")+
%  # theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(NZDep, perCap, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
%   #theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tp, aes(Age, perCap, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
%   #theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
tpsub<-tp[!(tp$Ethnicity=="MLA"),]
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tpsub, aes(NZDep, perCap, fill=Age)) + 
%   geom_bar(stat="identity", position="dodge")+
%   #theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tpsub, aes(NZDep, perCap, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
%   #theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% ggplot(tpsub, aes(Age, perCap, fill=Ethnicity)) + 
%   geom_bar(stat="identity", position="dodge")+
%   #theme(text=element_text(size=45))+
%   ylab("Per capita per 10000")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
# This chunk uses "results=tex"
library(Hmisc)
latex(PerCap, file="", table.env=FALSE,rowname=NULL)

## plot this figure?
#hist(tp$cases,xlab="Cases",main='Histogram of cases per category',col='grey',breaks=20)
tp$Ethnicity<- relevel(tp$Ethnicity, "European")
model<-glm(cases~Age*Ethnicity*NZDep+offset(log(Popn)),data=tp,family="quasipoisson")
summary(model)

model2<-update(model1,~.-Age:NZDep)
summary(model2)
anova(model2,test="F")

## use tpsub
#hist(tpsub$cases,xlab="Cases",main='Histogram of cases per category',col="darkgrey",breaks=20)

#model2<-update(model1,~.-Ethnicity:NZDep)
#summary(model2)
#anova(model2,test="F")
model2<-update(model1,~.-Age:NZDep)
summary(model2)
anova(model2,test="F")

model3<-update(model2,~.-Ethnicity:NZDep)
summary(model3)
anova(model3,test="F")
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(cex.axis=2)
% par(mar=c(5,6,4,2)+0.1)
% hist(tpsub$cases,xlab="Number of cases",main='',col='grey',breaks=20,ylab="Number of categories",cex.lab=2)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
res<-predict(model3)
#res<-predict(model2)
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% plot(exp(res),tpsub$cases,xlab="cases",ylab='predictions',main="",pch=16,col="darkgrey",cex.lab=2,cex=2)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% #hist(model3$residuals,main="",xlab="residuals",col="grey",cex.lab=2,ylab="")
% hist(model2$residuals,main="",xlab="residuals",col="grey",cex.lab=2,ylab="")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
library(xtable)
resan<-xtable(anova(model3,test="F"))
#resan<-xtable(anova(model2,test="F"))
print(resan,floating = FALSE)

library(xtable)
#xtb<-xtable(summary(model3))
xtb<-xtable(summary(model2))
print(xtb,floating = FALSE)

AgeVac<-table(testv$VC,testv$AgeInYears)
row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% barplot(t(AgeVac),xlab="Age",col=c("darkgrey","red","orange"),main="",ylab="Frequency",cex.lab=2)
% legend("topright",c("Unvaccinated","Dose 1","Dose 2"),fill=c("darkgrey","red","orange"),cex=1.8,bty="n")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=2)
% hist(vac$Dose1Mths,breaks=100,col=rgb(0,0,1,1/4),xlab="Age in months",main="",cex.lab=2,ylab="Frequency")
% abline(v=12,col="blue")
% hist(vac$Dose2Mths,breaks=50,add=T,col=rgb(1,0,0,1/4))
% abline(v=60,col="red")
% legend("topright",c("Dose 1","Dose 2"),col=c(rgb(0,0,1,1/4),rgb(1,0,0,1/4)),pch=15,bty="n",cex=1.8)
% legend("top",c("12 months","60 months"),col=c("blue","red"),lty=1,bty="n",cex=1.8)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=
DOB<-as.numeric(levels(testv$RptYear))[testv$RptYear]-testv$AgeInYears
#hist(DOB,breaks=100,col="grey",main="2007- 2014 measles cases",xlab="Date of Birth")

DOBVac<-table(testv$VC,DOB)
row.names(DOBVac)<-c("Unvaccinated","Dose1","Dose2")
DOBVac<-t(DOBVac)
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=1)
% barplot(t(DOBVac),xlab="Age",col=c("darkgrey","red","orange"),main="",ylab="Frequency",cex.lab=1)
% legend("topleft",c("Unvaccinated","Dose 1","Dose 2"),fill=c("darkgrey","red","orange"),cex=1,bty="n")
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=1)
% par(mfrow=c(2,1))
% barplot(DOBVac[,2],col=rgb(0,0,1,1/4),xlab="Date of Birth",main="",cex.lab=1,ylab="Frequency")
% legend("topleft",c("Dose 1"),col=rgb(0,0,1,1/4),pch=15,bty="n",cex=1)
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=1)
% barplot(DOBVac[,3],col=rgb(1,0,0,1/4),xlab="Date of Birth",main="",cex.lab=1,ylab="Frequency")
% legend("topleft",c("Dose 2"),col=rgb(1,0,0,1/4),pch=15,bty="n",cex=1)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}
% 
% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(mar=c(5,6,4,2)+0.1)
% par(cex.axis=1.5)
% par(mfrow=c(1,1))
% barplot(DOBVac[,1],col="grey",xlab="Date of Birth",main="",cex.lab=1.5,ylab="Frequency")
% legend("topleft",c("Unvaccinated"),col="grey",pch=15,bty="n",cex=1.5)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=

require(reshape2)  # this is the library that lets one flatten out data
require(ggplot2)   # plotting library
popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
@

% \begin{figure}
% \begin{center}
% <<fig=TRUE,echo=FALSE,width = 15,height=12>>=
% par(cex.axis=2)
% par(mar=c(5,6,4,2)+0.1)
% plot(pop,xlab="Age",ylab="Population",cex.lab=2)
% points(naive,pch=16)
% legend("topright",c("Population","Naive"),pch=c(1,16),bty="n",cex=1.8)
% @
% \end{center}
% \caption{test}
% %\label{fig:test}
% \end{figure}

<<echo=FALSE,include=FALSE,results=hide>>=

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
#naive[is.na(naive)] <- 0
colnames(naive)<-c("AgeInYears","Naive","Cases")
##

# proportions of the naive population the wuld be covered if vaccinated by age groups now
propvCohort<-c(
round((sum(naive$Naive[3:6])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:12])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:17])/sum(naive$Naive)),2),
round((sum(naive$Naive[6:19])/sum(naive$Naive)),2),
round((sum(naive$Naive[12:17])/sum(naive$Naive)),2),
round((sum(naive$Naive[12:19])/sum(naive$Naive)),2),
round((sum(naive$Naive[1:30])/sum(naive$Naive)),2),
round((sum(naive$Naive[1:45])/sum(naive$Naive)),2),
round((sum(naive$Naive[8:100])/sum(naive$Naive)),2))
names(propvCohort)<-c("2-5","5-11","5-16","5-18","11-16","11-18","0-29","0-44",">8")
write.table(propvCohort,"vaccine_proportions_Cohort.txt",sep=",",row.names=T,col.names=F)

propsv = NULL
for (i in 1:20 ) {
  propsv[i] = round((sum(naive$Naive[i:13])/sum(naive$Naive)),2)
}
propsvAge = NULL
for (i in 1:20 ) {
  propsvAge[i]=paste(c(min(i:20)-1, max(i:20)-1),collapse="-")
}

table120<-as.data.frame(propsv,propsvAge)
table120<-cbind(Age=rownames(table120),table120)
rownames(table120)<-1:20
colnames(table120)<-c("Age","Proportion")

propsvrev = NULL
for (i in 20:1 ) {
  propsvrev[i] = round((sum(naive$Naive[i:1])/sum(naive$Naive)),2)
}
propsvrevAge = NULL
for (i in 20:1 ) {
  propsvrevAge[i]=paste(c(min(i:1)-1, max(i:1)-1),collapse="-")
}

table201<-as.data.frame(propsvrev,propsvrevAge)
table201<-cbind(Age=rownames(table201),table201)
rownames(table201)<-1:20
colnames(table201)<-c("Age","Proportion")

propsv11 = NULL
for (i in 1:11 ) {
  propsv11[i] = round((sum(naive$Naive[i:11])/sum(naive$Naive)),2)
}
propsv11Age = NULL
for (i in 1:11 ) {
  propsv11Age[i]=paste(c(min(i:11)-1, max(i:11)-1),collapse="-")
}

table11<-as.data.frame(propsv11,propsv11Age)
table11<-cbind(Age=rownames(table11),table11)
rownames(table11)<-1:11
colnames(table11)<-c("Age","Proportion")

write.table(table11,"vaccine_proportions_11yrOlds.txt",sep=",",row.names=F)
write.table(table120,"vaccine_proportions_1_20yrOlds.txt",sep=",",row.names=F)
write.table(table201,"vaccine_proportions_20_1yrOlds.txt",sep=",",row.names=F)

###
@

@@

% \vspace{5mm} %5mm vertical space
% \begin{table}
% % <<label=regmodel,include=FALSE,results=tex,echo=FALSE>>=
% % library(xtable)
% % xtb<-xtable(summary(model3))
% % print(xtb,floating = FALSE)
% % @
% \caption{Summary of the regression model results for measles risk factors for cases from 2007-2014}
% \label{table:regression_riskfactors}
% \end{table}
% \vspace{5mm} %5mm vertical space
